---
title: "Simulations"
author: "Lauren Quesada"
date: "10/22/2021"
output: pdf_document
---
```{r}
###Load packages

library(tidyverse)
library(broom)

###Define function

LFunky<- function(beta0,beta1,beta2,n_obs,n_data,n_shuffle) {
  
  ###Define variables
  
  set.seed(4747)
  sig<- 0.05
  
  
    #If there is no intercept in the fitted regression, d.o.f. = (n-1)
    #otherwise, df=n-k-1; k=the number of variables
  dof <- n_obs-1
  
    #addp_val<-c()

    #intp_val<-c()
  
  simpp_val<-c()
  
  simp_t<-c()
  
  perm_pval<-c()
  
  ###Begin loop

  for(i in 1:n_data) {               # Start of the Monte Carlo loop

  ###Define Variables   
    
  x1<- rnorm(n_obs)     # Artificial x series, created just once

  x2<- x1+ rnorm(n_obs)     ## Did we figure out why these need to be defined in the loop?

  y<- beta0 + beta1*x1 + beta2*x2 + rnorm(n_obs)   # The DGP includes an intercept
  
  ###Fit a linear regression model
  
  simpfit<- lm(y ~ x1  -1)
  
  ###Record slope and p-value
  
  #observed p.value for each dataset (n_data)
  simpp_val<- c(simpp_val,
                simpfit %>% tidy() %>% filter(term =="x1") %>% select(p.value) %>% pull())
  
  #observed slope for each dataset (n_data)
  simp_t<- c(simp_t,
                simpfit %>% tidy() %>% filter(term =="x1") %>% select(statistic) %>% pull())
  
  #obs value for most recent dataset
  obsval<- simpfit %>% tidy() %>% filter(term =="x1") %>% select(statistic) %>% pull()
  
  #reassign vector so it doesn't add onto the concatenation each i in 1:n_data loop
  simpp_valperm<-c()
  
    ###Begin another loop
  
    for(j in 1: n_shuffle) {
  # A mis-specified model is estimated (unless Beta0 = 0)
    #addfit<- lm(y ~ x1+x2  -1) 
  
    #intfit<- lm(y ~ x1*x2  -1)
  
      
    ###Fit linear regression model to PERMUTED data
      
    simpfitperm<- lm(y ~ sample(x1)  -1)

    #addp_val<- c(addp_val, unname(coef(summary(addfit))[, "Pr(>|t|)"]))
  
    #intp_val<- c(intp_val, unname(coef(summary(intfit))[, "Pr(>|t|)"]))
    
    ###Record slope 
    simpp_valperm<- c(simpp_valperm,
                      simpfitperm %>% tidy() %>% filter(term == "sample(x1)") %>% select(statistic) %>% pull())

    }
  
  #p-value for each of my 200 datasets as compared to my permuted distribution of p-values
  perm_pval<-c(perm_pval,mean(simpp_valperm>obsval))
  
  }  #End of the Monte Carlo Loop

#coefficients, t-test p.values, permutation p.values
data.frame(simp_t, simpp_val, perm_pval)



  
}

#hist(addp_val, main="Distribution of  *ADDITIVE* p-Values", xlab="p-Value", freq=FALSE, border="black", col="yellow")

#hist(intp_val, main="Distribution of *INTERACTIVE* p-Values", xlab="p-Value", freq=FALSE, border="black", col="orange")

#hist(simpp_val, main="Distribution of *SIMPLE* p-Values", xlab="p-Value", freq=FALSE, border="black", col="red")

#results %>% ggplot() + geom_hist(aes(x = simppval))
```
In the function above, I am repeatedly (100 times) sampling 15 values/observations from a normal distribution. I have 15 ($x_i,y_i$) values in each dataset. I have 100 of these datasets. I am fitting a simple linear regression to each dataset, knowing I am not including my $x_2$ variable in the fit, knowing my data has two explanatory variables $x_1$ *and* $x_2$. Note: $x_2~x_1+rnorm(n_obs)$; $x_2$ is related to $x_1$, plus some noise. We *know* it is dependent. 
<<<<<<< HEAD

I am asking it to then record the observed slope ($\beta_1$) of this fitted line, as well as the p-value. Is there a correlation between my x and y variables ("no" implies $\beta_1=0$)? Permutation tests *assume* the null hypothesis to be true. We don't have any correlation between our data, and permuting should not change this. 

Next, I ask it to shuffle/permute my $x_1$ values, 200 times. Each dataset is shuffled 200 times, and each of those shuffles repeats the process described above. A line is fit to the permuted data, and the slope and p-value are recorded. Pretty soon, I have a distribution of permuted p-values and slopes.

But, I want to be able to compare my observed values to my permuted values, so I need the same size data frame for both. I take one of my 100 datasets. I shuffled it 200 times. I get a distribution of 200 permuted p-values for the 200 lines I fit (one for each permutation). I average the slopes of my lines to get 100 statistics. Then, I compare. How many permuted p-values are greater than my observed value? This is how I obtain my 100 p-values from my permuted data. If I shuffle my data, is there a correlation between my x and y variables?  

Remember, we've assumed the null hypothesis to be true. We're conducted this entire test assuming there is no correlation between my variables. So, which method is better at identifying this from the data? How do my normally distributed datasets compare to my permuted data--which is wrong *less* of the time?


**Should observe: Power(perm test) > Power(t-test)**

Power: By fixing our type I error, we’ve maximized the power of ourtest (5%), or maximized the probability of rejecting our nullhypothesis when it is false.

Power=1-P(Type I error)

# Fitted model, $H_o$ is true
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```
# Mis-Specified model, $H_o$ is true
***QUESTION*** Is this mis-specified if we subtract 1 to force it through the origin?
Answer: we "under-fit" the regression by omitting the intercept. The regression is fit through the origin, so unless the value of $\beta_1$ in our function is zero, the estimated model would be mis-specified:
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(1,0,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

=======

I am asking it to then record the observed slope ($\beta_1$) of this fitted line, as well as the p-value. Is there a correlation between my x and y variables ("no" implies $\beta_1=0$)? Permutation tests *assume* the null hypothesis to be true. We don't have any correlation between our data, and permuting should not change this. 

Next, I ask it to shuffle/permute my $x_1$ values, 200 times. Each dataset is shuffled 200 times, and each of those shuffles repeats the process described above. A line is fit to the permuted data, and the slope and p-value are recorded. Pretty soon, I have a distribution of permuted p-values and slopes.

But, I want to be able to compare my observed values to my permuted values, so I need the same size data frame for both. I take one of my 100 datasets. I shuffled it 200 times. I get a distribution of 200 permuted p-values for the 200 lines I fit (one for each permutation). I average the slopes of my lines to get 100 statistics. Then, I compare. How many permuted p-values are greater than my observed value? This is how I obtain my 100 p-values from my permuted data. If I shuffle my data, is there a correlation between my x and y variables?  

Remember, we've assumed the null hypothesis to be true. We're conducted this entire test assuming there is no correlation between my variables. So, which method is better at identifying this from the data? How do my normally distributed datasets compare to my permuted data--which is wrong *less* of the time?


>>>>>>> 36a67e54d1f0028b3d360de1077080854924b902
$Y_i=0+\beta_1*x_1+0*x_2$
$Y_i=\beta_1*x_1$
n=15, normally distributed
100 datasets
200 permutations
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

They're both right! The majority of p-values confirm there is *no* correlation between my variables ($\beta_1=0$), *BUT* my permuted p-values are more confidently saying this. More often and with a lower significance level, there are few permuted p-values larger than my observed value. 

Let's try increasing the slope, $\beta_1$ to 50 and see if this still holds true. I assume it will.

$Y_i=0+\beta_1*x_1+0*x_2$
$Y_i=\beta_1*x_1$

***QUESTION***: Do I need to change my intercept to -50 to force it through the origin? I get an error when I do this, why? Can I remove this from my function?
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,50,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```
What happened? I introduced a highly correlated slope, and my simple linear model was very good at recognizing the lack of correlation. But, it seems my permuted data failed *because* my simple linear model was so good. With a p-value of nearly zero in almost all of my 100 data sets, there was almost *never* a chance for the permuted values to be greater than my observed value? I'm not sure, but my p-values can't be negative. Is there some change that needs to be made in the function to avoid this?


Let's try a *smaller* slope,$\beta_1=0.1$ and see if this still holds true. I assume it will.

$Y_i=0+\beta_1*x_1+0*x_2$
$Y_i=\beta_1*x_1$
<<<<<<< HEAD

$\beta_1=0.1$
=======
>>>>>>> 36a67e54d1f0028b3d360de1077080854924b902
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.1,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```
Wrong again; with a slope very close to zero, I almost get a uniform distribution for both p-values. I assume if I ran more permutations, I would find a more uniform distribution. 
***QUESTION*** : I don't think uniform is the right word for the distribution since it's not [0,1], but what is?
Since the slope is so close to zero, all p-values are valid. Some data sets have correlation, some don't. I think my data is almost completely randomly distributed here, so p-values for both return a lack of consistency or conclusion. While it seems the simple p-values are more consistent across their distribution, the permuted p-values have a slightly higher frequency of low proportions. I wonder what would happen if I increased the number of permutations?

Let's try it!


<<<<<<< HEAD
Okay, let's put $\beta_1$ back to 1, but let's try a different intercept. How about $\beta_0=1$?
### I'm going to come back to this, but first I want to try to reproduce David Giles' results where he manipulates $\beta_1$

**For n=15**
$\beta_1=0.2$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.2,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.3$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.3,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.4$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.4,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.5$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.5,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.6$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.6,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.7$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.7,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.8$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.8,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.9$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.9,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.1$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.1,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.2$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.2,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.3$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.3,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.4$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.4,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.5$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.5,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.6$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.6,0,15,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```


***QUESTION*** Do I have to name all these differently in order to plot them/how do I make his power plots? 

**Should observe: Power(perm test) > Power(t-test)**
David Giles doesn't test past $\beta_1=1$, at which point it seems like Power(perm test)=Power(t-test). What do we observe when $\beta_1>1$?

**For n=30**
# Fitted model, $H_o$ is true
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

# Mis-Specified model, $H_o$ is true
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(1,0,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.1$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.1,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.2$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.2,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.3$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.3,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.4$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.4,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.5$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.5,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.6$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.6,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.7$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.7,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.8$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.8,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.9$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.9,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.1$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.1,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.2$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.2,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.3$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.3,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.4$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.4,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=1.5$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,1.5,0,30,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

**Should observe: Power(perm test) < Power(t-test)**
David Giles only goes up to $\beta_1=0.5$, for which the t-test has a power of 1, and the permutation test is approaching a power of 1. 

**For n=60**
# Fitted model, $H_o$ is true
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

# Mis-Specified model, $H_o$ is true
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(1,0,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.05$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.05,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.1$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.1,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.15$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.15,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.2$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.2,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.25$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.25,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.3$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.3,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.35$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.35,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.4$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.4,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.45$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.45,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=5$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.5,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.55$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.55,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.6$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.6,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.65$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.65,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```

$\beta_1=0.7$
```{r}
#beta0,beta1,beta2,n_obs,n_data,n_shuffle
results<-LFunky(0,0.7,0,60,100,200)
hist(results$simpp_val)
hist(results$perm_pval)
```


=======
Okay, let's put $\beta_1$ back to 1, but let's try a different intercept. How about $\beta_0=1$?
>>>>>>> 36a67e54d1f0028b3d360de1077080854924b902
